Conex - establishing trust into data repositories

### 2016-10-13 paper.tex
some updates

\begin{abstract}
  Opam is a software update systems, responsible for discovering, downloading, building, and installing packages.
  The data and metadata at each step should be authenticated to originate from a (transitively) trusted author.
  Opam does not include any mechanism to authenticate any step at the moment.
  We propose \emph{conex}, which establishes end-to-end digital signatures from the package author to the user.
  Using conex, neither the distribution server nor the transport layer need to be trusted.
  Authors and users need to adapt their workflow only slightly (or use opam-publish), whereas janitors (repository maintainers) need more tooling.
\end{abstract}

\section{Introduction}
Opam discovers, downloads, builds, and installs OCaml packages.
To prevent injection of malware, all received data needs to be authenticated.
A software update system relies on a central server which distributes a repository (package metadata: dependencies and build instructions) of packages.
When a new package is released, it is inserted into the repository and distributed to all clients by the central server.
The central server is an online service and a single point of failure.
Recently those servers have been under attack \cite{adobe,apache2,apache1,opera,kernel,cabal,kernelorg,debian,freebsd,php3,savannah,maven,savannah2,npm,php,phpmyadmin,php2,github,ocamlforge,redhatssh,gentoo,rubygems,fedora,wordpress} to inject malware into packages.

@@

In this paper we present \emph{conex}, a system which establishes end-to-end digital signatures - the original author signs their package release and build instructions, the client verifies this signature - to re-establish trust into packages installed by opam.
Using end-to-end signatures, neither the the server hosting the repository nor the server where the tarballs are downloaded from need to be trusted!
Conex has been designed with opam in mind, but it can be used without opam to publish and update any data in an authenticated way.
Conex is not limited to the main opam repository, but can be used with any public or private opam repository, or any other data repository.
The only requirement is a second channel to distribute an initial set of trust anchors (e.g. an opam package in the main opam repository).
The main goal of conex is to \emph{establish an authenticated way to distribute packages}.

@@

Conex is implemented in OCaml and licensed under the 2-clause BSD license (source~\cite{conex}).
To provide authenticity from the beginning, conex is best installed together with opam.

Earlier work on securing software update systems, such as the update framework (TUF)~\cite{tuf} and Diplomat~\cite{diplomat} focus on distributing directories securely, but due to centralised files simultaneous updates of packages are not free of conflicts.
Our earlier proposal for opam signing~\cite{opamsigning} introduced unnecessary structure and complexity by adjusting TUF to the decentralised opam workflow.

In section~\ref{sec:design} we discuss the design of conex, followed by reasonable system optimisations in section~\ref{sec:shortcuts}.
Afterwards in section~\ref{sec:implementation} we briefly present the implementation of conex and opam integration.
Related work is mentioned in section~\ref{sec:related}, we conclude in section~\ref{sec:conclusion}.

@@

Conex cannot detect compromises of (offline!) keys.  But if janitors, authors, and other people carefully read through updates of packages, surprising releases will be detected.
We assume that authors and janitors protect their private keys on their computers in a reasonable way, but we also have mechanisms in place for key revocation.
If a private key is compromised, we consider conex security to be effective if the impact is limited to the authorised packages of the compromised key.

\subsection{Roles}
There are two roles in conex:
\begin{itemize}
  \item An \emph{author} who develops packages and releases them to the repository.  Each author signs their own packages.
  \item A \emph{janitor} who keeps the repository tidy and in a working state (fix up reverse dependencies etc.).  Janitors also maintain the package name and author name resources by signing them.  Authorisations, that an author is responsible for a package, are signed by janitors as well.  Instead of trusting a single janitor to sign these resources, a quorum of signatures of janitors is required to avoid this concentration of power.
\end{itemize}

A janitor may also act as an author, in case they develop and maintain packages on their own.  There is no need for a janitor to have multiple private keys.

\subsection{Resources}
We distinguish various resources which are all hosted in the same repository:
author information (public key, id), teams (members), package metadata (build recipe, dependencies, url to tarball), list of releases (of a single package), and authorisation (relation between package name and author).
The repository contains one index per author with all resources and checksums the author vouches for.
A janitor is a member of the team ``janitors''.

Each resource is only valid if it is signed by either one of the authorised key identifiers (in case of relesaes and package metadata checksums), or a quorum of janitors.
Each resource contains a monotonic counter to prevent rollback attacks.

@@

Each author updates their own packages, which should be free of conflicts between concurrent repository updates, due to separate resources for each package.
% unclear without VC in mind, which is'nt introduced yet: precise ``conflicts between concurrent repository updates'' ?
If a conflict is detected, it is up to the janitors to decide which update should go through, and which should be resubmitted on top of the other.
The main source of conflict is likely team membership addition and deletion.

\subsection{Repository layout}
Conex is not bound to the file system, but we use their terminology.
The root directory has three subdirectories, \texttt{keys}, \texttt{index} and \texttt{packages}.

The \texttt{keys} directory contains one file for each author and each team, named after their identifier.
The \texttt{index} directory contains one index file for each author, named after the key identifier.  This contains a list of resources the author vouches for, and a digital signature.
The \texttt{packages} directory consists of subdirectories, which each must contain a file \texttt{authorisation} and a file \texttt{releases}, plus possibly any subdirectories.
Each subdirectory must be listed in the \texttt{releases} file, and must contain a \texttt{checksum} file besides any other data files and directories (all listed in the checksum file).

Each file contains a set of key-value pairs.
Any key-value representation which can be non-ambiguously normalised into a string, can be used (such as the opam format, json, s-expressions).
Each index file contains a \texttt{signed} structure: \texttt{(signed: }\textbf{resource}\texttt{, sigs: }\textbf{(id, signature) list}\texttt{)}.
The amount of public key operations for verification is at most the number of authors of the repository.
The different resources are explained in more detail below.
The signature is computed over the normalised string representation of the resource concatenated with the public key id, separated by a space character.

\begin{figure}[h]
\textbf{id}: restricted string\\
\textbf{name}: restricted string\\
\textbf{signature}: Base64 encoded RSASSA-PSS with SHA256 (PKCS1)\\
\textbf{public key}: RSA public key ($\geq$ 2048 bit, PKCS1, PEM encoded)\\
\textbf{service}: \texttt{Email} of \textbf{name} $|$ \texttt{GitHub} of \textbf{name} $|$ \texttt{Other} of \textbf{id} * \textbf{name}\\
\textbf{digest}: SHA256 digest of file\\
\textbf{byte size}: size as \textbf{Int64} of file
\caption{\label{fig:datatypes0}
  Data types for conex version 0}
\end{figure}

@@

Each file includes a data version (0 in this proposal) and a monotonic 64 bit counter (starting at 0).

\begin{figure}[h]
  \begin{minipage}{\hsize}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
key: \textbf{public key},\\
identifier: \textbf{id},\\
accounts: \textbf{service list},\\
 \caption{\label{fig:keyformat} Public key}
    \end{minipage}
    \hspace*{.5cm}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{id},\\
members: \textbf{id list}
 \caption{\label{fig:teamformat} Team}
    \end{minipage}
  \end{minipage}
\end{figure}

@@

\begin{figure}[h]
  \begin{minipage}{\hsize}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{name},\\
authorised: \textbf{id list}
 \caption{\label{fig:authformat} Authorisation}
    \end{minipage}
    \hspace*{.5cm}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{name},\\
releases: \textbf{name list}
 \caption{\label{fig:releasesformat} Releases}
    \end{minipage}
  \end{minipage}
\end{figure}

@@

\begin{figure}[h]
  \begin{minipage}{\hsize}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{name},\\
files: \textbf{(name, size, digest) list}
 \caption{\label{fig:checksumformat} Checksum}
    \end{minipage}
    \hspace*{.5cm}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
identifier: \textbf{id},\\
resources: \textbf{(name, kind, digest) list}
 \caption{\label{fig:janindexformat} Index}
    \end{minipage}
  \end{minipage}
\end{figure}

@@

\paragraph{Update verification}  The repository on disk is already verified.
The client uses it to verify each update separately.
Each modified resource must be signed, either by an authorised author or by a quorum of janitors.
The counter must start at 0 (upon creation) and increment.
The file name and identifier must match.
A key rollover needs to introduce the new key (replacing the old), and sign again all resources which were signed with the old key.
Removal of resources is done by replacing it with the empty string.
A new package needs to have an authorisation file and a releases file, both properly signed.
%% \begin{itemize}
%%   \item janitor index modification
%%   \item janitor index creation - counter 0, properly named, signed by janitor
%%   \item janitor index deletion - counter increasing, properly named, signed by quorum
%%   \item key modification -- key is sufficiently big, unique id matching filename, counter is increased, signed by quorum, no signature with old key
%%   \item key insertion -- key is sufficiently big, unique id matching filename, counter is 0, signed by quorum
%%   \item key deletion -- counter increasing, no signatures with this key, empty string is used to denote no key, signed by quorum
%%   \item role modification -- counter increasing, no signature with old role, signed by quorum
%%   \item authorisation insertion -- counter is 0, id matches directory, signed by quorum, authorised key ids are present
%%   \item authorisation modification -- counter increasing, all checksum are signed with an authorised key, signed by quorum
%%   \item complete package removal -- empty authorised, empty releases, signed by quorum
%%   \item package insertion -- valid authorisation, releases, and checksum, both counter 0, signed by author (authorisation by quorum)
%%   \item package update -- valid checksum, counter increasing, signed by author
%%   \item new package version release -- index counter increasing and new version added to list, valid checksum, signed by author
%%   \item package version deletion -- counter increasing, removed from releases, signed by author
%% \end{itemize}

@@

Each resource includes a monotonically increasing counter to prevent rollback attacks.
Otherwise, an attacker could replay e.g. an insecure package build recipe.
Also, each resource includes a name, otherwise the signature of package A would be valid for package B.
Each signature contains the resource kind to avoid carrying over a signature from one resource to another.

Instead of removing resources to revoke them, the already claimed names stay there.
The public key, package, or janitor index resource is emptied and must be signed by a quorum of janitors.

A person-in-the-middle attacker could implement a backdoor for a package, but would need to either modify the checksum, prepare a new release, embed the patch as metadata in the repository, or introduce a new package (which they could) and add a dependency.
All of these actions require either a signature from the original author or from a quorum of janitors.
Any suspicious behaviour in widely used packages will likely be spotted by the community which reads through the commit logs.

@@

Mix-and-match attacks provide e.g. outdated releases of package A and B, up-to-date releases of package C, are not mitigated in conex since there is no central list of releases.
They can be mitigated by relying on some append-only log, blockchain, or other monotonic updates, such as git without forced pushes.
We will discuss a timestamp notary in the next section, mitigating both mix-and-match and freeze attacks.

\section{Optimisations} \label{sec:shortcuts}
%\TODO{see also http://bentnib.org/posts/2016-04-12-authenticated-data-structures-as-a-library.html}

@@

To mitigate mix-and-match attacks, a \emph{timestamp notary}, which updates the repository regularly, verifies all updates, and attaches its digital signature to the latest patch (via an git annotated tag over the git commit).
The public key of the timestamp notary is inside the repository and signed by multiple janitors.
Clients can, instead of verifying packages and updates independently, trust the timestamp notary (or multiple timestamp notaries).
This will reduce computational overhead (signature verification are expensive operations on big numbers), and might be worth for embedded devices.
If there are multiple instances of a timestamp notary run by different organisations, and its code is well audited, it might be reasonable to use them as default.
While there is not yet an implementation of a timestamp notary, it will likely be implemented as a MirageOS~\cite{mirage,nqsb-tls} application, thus the trusted code base is small.

The main opam repository served as an example over the last section, but conex can easily be applied to any other opam repository.
Initial set of janitors (can be a single element), the quorum (can be one) must be distributed for each other repository.
Of course, an automated timestamp notary can be used for other repositories as well.

%\TODO{this is insecure, since an attacker can add themselves to the authorisation file and self-sign it!}
If the work overhead for janitors turns out too high, we can leverage the system in a way that \emph{neither authorisation nor public keys require a quorum}.
This will open the system to name squatting attacks.
Snapshot verification does not apply anymore, because there is no history of authorisations.
Update verification can still be done, thus we could trust the timestamp notaries and use git.

%that they contain the same information as the original ones, but updates can be verified (and again, we can outsource trust into the timestamping notary).

@@

Conex is work-in-progress~\cite{conex}.
It will be abstracted over the data layout (json, opam, s-expression), and over a cryptography provider.
One provider will be nocrypto, another the \emph{openssl} utility, enabling bootstrapping without opam.
The timestamp notary and other automated tools will use the conex library.

@@

More recent is Diplomat~\cite{diplomat}, which automatically promoting unclaimed packages to claimed (which we can do using timestamp notary).
Our security model instead uses end-to-end signatures of packages, and relies on authors to sign their packages.

@@

Our original opam signing proposal~\cite{opamsigning} did not include a list of releases per package, opening up to various attacks.
Additionally, our original proposal introduced an unnecessary hierarchical structure.
It relied on the timestamping notary or using git as update protocol with appropriate protection of the transport layer.

@@

Future work includes finishing the implementation of conex, integrating it into opam, and developing automated services (timestamp notary, build bot, PR bot).
A first beta version of conex, including opam integration, is expected to be available for the OCaml workshop 2016.

>>>>

\begin{abstract}
  Opam is a software update system, responsible for discovering, downloading, building, and installing packages from community repositories.
  The data and metadata at each step should be authenticated to originate from a (transitively) trusted author.
  Opam does not include any mechanism to authenticate any step at the moment.
  We propose \emph{conex}, which establishes digital signatures from the package author to the user.
  Using conex, neither the distribution server nor the transport layer need to be trusted.
  Clients will install authenticated packages, and notice conex only if signatures cannot be verified.
  Authors need to adapt their workflow slightly, an updated \emph{opam publish} will be developed.
  Janitors (repository maintainers) need to adapt their workflow to emit cryptographic signatures to the repository (in addition to pressing merge PR buttons).
\end{abstract}

\section{Introduction}
Opam discovers, downloads, builds, and installs OCaml packages from community repositories.
It is a software update system which relies on a central server which distributes a repository (package metadata: dependencies and build instructions) of packages.
When a new package is released, it is inserted into the repository and distributed via the central server to all clients pulling it.
The central server is an online service and a single point of failure.
Recently those servers have been under attack \cite{adobe,apache2,apache1,opera,kernel,cabal,kernelorg,debian,freebsd,php3,savannah,maven,savannah2,npm,php,phpmyadmin,php2,github,ocamlforge,redhatssh,gentoo,rubygems,fedora,wordpress} to inject malware into packages.

@@

In this paper we present \emph{conex}, a system which establishes digital signatures - the original author signs their package release and build instructions, the client verifies this signature - to re-establish trust into packages installed by opam.
Neither the the server hosting the repository nor the server where the tarballs are downloaded from need to be trusted!
Conex has been designed with opam in mind, but it can be used without opam to publish and update any data in an authenticated way.
Conex is not limited to the main opam repository, it can be used with any public or private opam repository.
The only requirement is a second channel to distribute an initial set of trust anchors (e.g. an opam package in the main opam repository).
The main goal of conex is to \emph{establish an authenticated way to distribute packages}.

@@

Conex is implemented in OCaml and licensed under the 2-clause BSD license (source~\cite{conex}).
To provide authenticity for all other packages, conex is best installed together with opam.
This leads to a constraint on conex: it may not have any dependencies apart from opam and the OCaml runtime.
The installation of OCaml, opam, and conex need to be authenticated via other means (such as the OS package manager).

Earlier work on securing software update systems, such as the update framework (TUF)~\cite{tuf} focusses on distributing directories securely, but due to centralised files simultaneous updates of packages are not free of conflicts.
Diplomat~\cite{diplomat} was designed at a similar time as conex, and shares various goals as conex.
%XXX: there need to be a more concrete comparison with diplomat somewhere
Our earlier proposal for opam signing~\cite{opamsigning} introduced unnecessary structure and complexity by adjusting TUF to the decentralised opam workflow.

In section~\ref{sec:design} we discuss the design of conex, followed by a timestamping service in section~\ref{sec:shortcuts}.
Afterwards in section~\ref{sec:implementation} we briefly present the implementation of conex and opam integration.
Related work is mentioned in section~\ref{sec:related}, we conclude in section~\ref{sec:conclusion}.

@@

%(offline!) keys -- we need to be precise which keys are where..
Conex cannot detect compromises of keys.  But if janitors, authors, and other people carefully read through updates of packages, releases not done by the original author will be detected.
We assume that authors and janitors protect their private keys on their computers in a reasonable way, but we also have mechanisms in place for key revocation.
If a private key is compromised, we consider conex security to be effective if the impact is limited to the authorised packages of the compromised key.

\subsection{Roles}
There are two roles in conex:
\begin{itemize}
  \item An \emph{author} who develops packages and releases them to the repository.  Each author signs their own packages.
  \item A \emph{janitor} who keeps the repository in a working state (fix up reverse dependency constraints etc.).  Janitors also maintain the package name and author name resources by signing them.  Instead of trusting a single janitor to sign these resources, a quorum of signatures of janitors is required to avoid this concentration of power.
\end{itemize}

A janitor may also act as an author, in case they develop and maintain packages on their own.  There is no need for a janitor to have multiple private keys.
The set of janitors is a subset of the authors.

\subsection{Resources}
We distinguish various resources which are all hosted in the same repository:
author information (public key, id, contact information), teams (members), package metadata (build recipe, dependencies, url to tarball), list of releases (of a single package), and authorisation (who is maintaining each package).
For each author one signed index file is in the repository, containing all resources and digests the author vouches for.
A janitor is a member of the team ``janitors''.

Each resource is only valid if it is vouched for either by an authorised key (in case of relesaes and package checksums), or a quorum of janitors.
Each resource contains a counter, monotonically increasing between updates, to prevent rollback attacks.
A quorum of janitors can reset the counter to recover from fast forward attacks.

@@

Each author updates their own packages, which should be free of conflicts between concurrent repository updates, due to separate resources for each package.
% unclear without VC in mind, which is'nt introduced yet: precise ``conflicts between concurrent repository updates'' ?
If a conflict is detected, it is up to the janitors to decide which update should go through, and which should be resubmitted on top of the other.
%Modifications to team membership can only be done sequentially 
%The main source of conflict is likely team membership addition and deletion.

\subsection{Repository layout}
Conex is not bound to a file system, but we use their terminology.
The root directory has three subdirectories, \texttt{keys}, \texttt{index} and \texttt{packages}.

The \texttt{keys} directory contains one file for each author and each team, named after their identifier.
The \texttt{index} directory contains one index file for each author, named after their key identifier.
The \texttt{packages} directory consists of subdirectories, which each must contain an \texttt{authorisation} file and a \texttt{releases} file, plus possibly any subdirectories.
Each subdirectory must be listed in the \texttt{releases} file, and must contain a \texttt{checksum} file besides any other data files and further subdirectories (all listed in the checksum file).

Each file contains a set of key-value pairs.
Any key-value representation which can be non-ambiguously normalised into a string, can be used (such as the opam format, json, s-expressions).
Each index file contains a \texttt{signed} structure: \texttt{(signed: }\textbf{resource}\texttt{, sigs: }\textbf{(id, signature, timestamp) list}\texttt{)}.
The amount of public key operations for verification is at most the number of authors of the repository.
The different resources are explained in more detail below.
The signature is computed over the normalised string representation of the resource concatenated with the public key id, and the current timestamp, using a space character as separator.

\begin{figure}[h]
\textbf{id}: restricted string\\
\textbf{name}: restricted string\\
\textbf{timestamp}: Seconds since UNIX epoch as \textbf{Int64}\\
\textbf{signature}: Base64 encoded RSASSA-PSS with SHA256 (PKCS1)\\
\textbf{public key}: RSA public key ($\geq$ 2048 bit, PKCS1, PEM encoded)\\
\textbf{service}: \texttt{Email} of \textbf{name} $|$ \texttt{GitHub} of \textbf{name} $|$ \texttt{Other} of \textbf{id} * \textbf{name}\\
\textbf{digest}: SHA256 digest of file\\
\textbf{byte size}: size as \textbf{Int64} of file\\
\textbf{resource type}: PublicKey $|$ Team $|$ Checksum $|$ Releases $|$ Authorisation
\caption{\label{fig:datatypes0}
  Data types for conex version 0}
\end{figure}

@@

Each file includes a version (0 for this proposal) and a monotonic 64 bit counter (starting at 0).

\begin{figure}[h]
  \begin{minipage}{\hsize}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
key: \textbf{public key},\\
identifier: \textbf{id},\\
accounts: \textbf{service list}
 \caption{\label{fig:keyformat} Public key file}
    \end{minipage}
    \hspace*{.5cm}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{id},\\
members: \textbf{id list}
 \caption{\label{fig:teamformat} Team file}
    \end{minipage}
  \end{minipage}
\end{figure}

@@

\begin{figure}[h]
  \begin{minipage}{\hsize}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{name},\\
authorised: \textbf{id list}
 \caption{\label{fig:authformat} Authorisation file}
    \end{minipage}
    \hspace*{.5cm}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{name},\\
releases: \textbf{name list}
 \caption{\label{fig:releasesformat} Releases file}
    \end{minipage}
  \end{minipage}
\end{figure}

@@

\begin{figure}[h]
  \begin{minipage}{\hsize}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{name},\\
files: \textbf{(name, size, digest) list}
 \caption{\label{fig:checksumformat} Checksum file}
    \end{minipage}
    \hspace*{.5cm}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
identifier: \textbf{id},\\
resources: \textbf{(name, resource type, digest) list}
 \caption{\label{fig:janindexformat} Index file}
    \end{minipage}
  \end{minipage}
\end{figure}

@@

\paragraph{Update verification}  The repository on disk is already verified.
The client uses it to verify each update separately.
Each modified resource must be signed, either by an authorised author or by a quorum of janitors.
The counter must start at 0 (upon creation) and increment.
The file name and identifier must match.
A key rollover needs to introduce the new key (replacing the old), and a signature with the new key of the index file.
Removal of resources is done by replacing it with the empty string.
A new package needs to have an authorisation file (signed by a quorum) and a releases file (signed by an authorised author).
%% \begin{itemize}
%%   \item janitor index modification
%%   \item janitor index creation - counter 0, properly named, signed by janitor
%%   \item janitor index deletion - counter increasing, properly named, signed by quorum
%%   \item key modification -- key is sufficiently big, unique id matching filename, counter is increased, signed by quorum, no signature with old key
%%   \item key insertion -- key is sufficiently big, unique id matching filename, counter is 0, signed by quorum
%%   \item key deletion -- counter increasing, no signatures with this key, empty string is used to denote no key, signed by quorum
%%   \item role modification -- counter increasing, no signature with old role, signed by quorum
%%   \item authorisation insertion -- counter is 0, id matches directory, signed by quorum, authorised key ids are present
%%   \item authorisation modification -- counter increasing, all checksum are signed with an authorised key, signed by quorum
%%   \item complete package removal -- empty authorised, empty releases, signed by quorum
%%   \item package insertion -- valid authorisation, releases, and checksum, both counter 0, signed by author (authorisation by quorum)
%%   \item package update -- valid checksum, counter increasing, signed by author
%%   \item new package version release -- index counter increasing and new version added to list, valid checksum, signed by author
%%   \item package version deletion -- counter increasing, removed from releases, signed by author
%% \end{itemize}

@@

Each resource includes a monotonically increasing counter to prevent rollback attacks.
Otherwise, an attacker could replay e.g. an insecure package build recipe.
To prevent fast forward attacks, a quorum of janitors can reset counters.
Also, each resource includes a name, otherwise the signature of package A would be valid for package B.
Each signature contains the resource type to avoid carrying over a signature from one resource to another.

Instead of removing resources to revoke them, the already claimed names stay there.
The public key or package is emptied and must be signed by a quorum of janitors.

A person-in-the-middle attacker could implement a backdoor for a package, but would need to either modify the checksum, prepare a new release, embed the patch as metadata in the repository, or introduce a new package and add a dependency.
All of these actions require either a signature from the original author or from a quorum of janitors.
Any suspicious behaviour in widely used packages will likely be spotted by the community which reads through the commit logs.

@@

Mix-and-match attacks provide e.g. outdated releases of package A and B, up-to-date releases of package C, are not mitigated in conex since there is no central list of releases.
They can be mitigated by relying on some append-only log, blockchain, or other monotonic updates, such as git without forced pushes.
We will discuss a timestamping service in the next section, mitigating both mix-and-match and freeze attacks.

\section{Timestamping Service} \label{sec:shortcuts}
%\TODO{see also http://bentnib.org/posts/2016-04-12-authenticated-data-structures-as-a-library.html}

@@

To mitigate mix-and-match attacks, a \emph{timestamp service}, which updates the repository regularly, verifies all updates, and attaches its digital signature to the latest patch.
The public key of the timestamp service is part of the repository and signed by multiple janitors.
If there are multiple instances of a timestamp service run by different organisations, and its code is well audited, it might be reasonable to only verify their signatures, and trust it to verify all author signatures.
While there is not yet an implementation of a timestamp service, it will likely be implemented as a MirageOS~\cite{mirage,nqsb-tls} application, thus the trusted code base is small.

The main opam repository served as an example over the last section, but conex can easily be applied to any other opam repository.
Initial set of janitors (can be a single element), the quorum (can be one) must be distributed for each other repository.
Of course, an automated timestamp service can be used for other repositories as well.

%\TODO{this is insecure, since an attacker can add themselves to the authorisation file and self-sign it!}
If the work overhead for janitors turns out too high, we can leverage the system in a way that \emph{neither authorisation nor public keys require a quorum of janitors}.
Diplomat~\cite{diplomat} introduced online keys for this (their \texttt{new} role).
This will open the system to name squatting attacks.
Snapshot verification does not apply anymore, because there is no history of authorisations.
Update verification can still be done, thus we could trust the timestamp notaries and use git.

%that they contain the same information as the original ones, but updates can be verified (and again, we can outsource trust into the timestamping service).

@@

Conex is work-in-progress~\cite{conex}.
It will be abstracted over the data layout (json, opam, s-expression), and over a cryptography provider.
One provider will be nocrypto, another the \emph{openssl} utility, enabling bootstrapping without opam.
The timestamp service and other automated tools will use the conex library.

@@

More recent is Diplomat~\cite{diplomat}, which automatically promoting unclaimed packages to claimed (which we can do using timestamp service).
Our security model instead uses digital signatures of packages, and relies on authors to sign their packages.

@@

Our original opam signing proposal~\cite{opamsigning} did not include a list of releases per package, opening up to various attacks.
Additionally, our original proposal introduced an unnecessary hierarchical structure.
It relied on the timestamping service or using git as update protocol with appropriate protection of the transport layer.

@@

Future work includes finishing the implementation of conex, integrating it into opam, and developing automated services (timestamp service, build bot, PR bot).
A first beta version of conex, including opam integration, is expected to be available for the OCaml workshop 2016.
